---
title: "Forecasting the Term Structure of Interest Rates"
subtitle: "Macroeconomics -- Assignment 2, Exercise 1"
author: "Milo Barbavara, Niccolò Bianchi"
date: "`r Sys.Date()`"
output:
  pdf_document:
    toc: true
    number_sections: true
    fig_caption: true
fontsize: 11pt
geometry: margin=2.5cm
header-includes:
  - \usepackage{booktabs}
---

```{r setup, include=FALSE}
# Global knitting options: hide code by default, suppress messages/warnings,
# centre figures. Individual chunks override 'echo' to show crucial snippets.
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE,
                      fig.align = "center", fig.width = 7, fig.height = 4.2)
library(fredr); library(xts); library(zoo); library(vars)
fredr_set_key("2dc39adca864462a604d64d7ea499289")   # personal FRED API key
lambda <- 0.0609                                     # Nelson-Siegel decay (fixed by assignment)
```

# Introduction and roadmap

The **yield curve** is the set of interest rates at which the US government can
borrow across different horizons (maturities), observed at a single point in
time. Because the rates at neighbouring maturities move almost together, the
entire curve -- eight maturities in our data -- can be summarised by just three
quantities: its overall **level**, its **slope** (long minus short), and its
**curvature** (the bend in the middle). This report uses that idea to *forecast*
the curve.

We follow two complementary econometric approaches:

1. a **two-step dynamic Nelson--Siegel** model, where the curve is first
   compressed each month into three factors (Question 2) and those factors are
   then forecast with a **VAR** (Question 3);
2. a **one-step state-space** model, where the same three factors are treated as
   unobserved states and estimated jointly with their dynamics by the
   **Kalman filter** (Question 4).

Question 1 prepares the data and motivates the three-factor structure
descriptively; Question 5 compares the models and draws the conclusions.

```{r data, include=FALSE, cache=TRUE}
# --- Q1 data construction -----------------------------------------------------
fred_ids   <- c("DGS3MO", "DGS6MO", "DGS1", "DGS2", "DGS3", "DGS5", "DGS7", "DGS10")
maturities <- c(3,        6,        12,     24,     36,     60,     84,     120)

daily_list <- list()
for (i in seq_along(fred_ids)) {
  raw <- fredr(series_id = fred_ids[i])
  daily_list[[i]] <- xts(raw$value, order.by = raw$date)
}
daily <- do.call(merge, daily_list)
colnames(daily) <- as.character(maturities)

monthly <- apply.monthly(daily, function(x) colMeans(x, na.rm = TRUE))   # daily -> monthly average
yields  <- na.omit(monthly)                                              # balanced panel
index(yields) <- as.Date(as.yearmon(index(yields)))

# Empirical (model-free) level, slope, curvature
emp_level     <- yields[, "120"]
emp_slope     <- yields[, "120"] - yields[, "3"]
emp_curvature <- 2 * yields[, "24"] - yields[, "3"] - yields[, "120"]
empirical <- merge(emp_level, emp_slope, emp_curvature)
colnames(empirical) <- c("Level", "Slope", "Curvature")
persist <- sapply(empirical, function(x) acf(x, lag.max = 1, plot = FALSE)$acf[2])
```

# Question 1 -- Data preparation and descriptive analysis

We download **daily** US Treasury constant-maturity yields from FRED for eight
maturities (3, 6, 12, 24, 36, 60, 84 and 120 months), chosen so that the panel
is *balanced* (no long gaps) and contains the 3-, 24- and 120-month maturities
required by the empirical formulas below. Two cleaning choices matter: we take
**monthly averages** of the daily data (to match the frequency of macro
variables and smooth daily noise), and we **drop months with any missing
maturity** (the models need a complete cross-section every month).

The resulting sample runs from **`r format(start(yields))`** to
**`r format(end(yields))`**, with **`r nrow(yields)`** monthly observations on
the eight maturities listed above.

## The shape of the curve at three dates

```{r q1-curves, fig.cap="Yield curve at an early, a middle and a recent date. An upward slope is the normal case; a flat or inverted curve (long below short) typically signals an expected slowdown."}
idx_dates <- c(1, round(nrow(yields) / 2), nrow(yields))
plot(maturities, as.numeric(yields[idx_dates[1], ]), type = "b", pch = 19,
     ylim = range(yields[idx_dates, ]), col = "darkblue",
     xlab = "Maturity (months)", ylab = "Yield (%)", main = "Yield curve at three dates")
lines(maturities, as.numeric(yields[idx_dates[2], ]), type = "b", pch = 17, col = "darkgreen")
lines(maturities, as.numeric(yields[idx_dates[3], ]), type = "b", pch = 15, col = "darkred")
legend("bottomright", bty = "n", pch = c(19, 17, 15),
       col = c("darkblue", "darkgreen", "darkred"),
       legend = format(index(yields)[idx_dates], "%Y-%m"))
```

**Interpretation.** The *shape* of the curve is itself a macroeconomic signal.
An upward-sloping curve reflects normal growth expectations (lending longer
earns more). A flat or **inverted** curve -- long rates at or below short rates
-- means markets expect the central bank to cut rates, which usually accompanies
an expected recession. The three dates show the curve in different rate regimes.

## Empirical level, slope and curvature

We summarise each monthly curve with three model-free quantities:
$$
\text{Level}_t = y_t(120), \qquad
\text{Slope}_t = y_t(120) - y_t(3), \qquad
\text{Curvature}_t = 2\,y_t(24) - y_t(3) - y_t(120).
$$

```{r q1-empirical, fig.height=5.6, fig.cap="Empirical level (long rate), slope (long minus short) and curvature (medium-term bend) over time."}
plot.zoo(empirical, main = "Empirical level, slope and curvature",
         xlab = "Date", col = c("darkblue", "darkred", "darkgreen"))
```

**Persistence and cyclical behaviour.** The first-order autocorrelations are
Level = `r round(persist["Level"],2)`, Slope = `r round(persist["Slope"],2)`,
Curvature = `r round(persist["Curvature"],2)`.

* The **level** is the most persistent series (autocorrelation near 1): it tracks
  long-run interest-rate and inflation regimes that shift only slowly, behaving
  almost like a random walk.
* The **slope** is persistent but clearly **cyclical**: it flattens or inverts
  ahead of recessions and steepens in recoveries, so it co-moves with the
  business cycle and reflects the stance of monetary policy.
* The **curvature** is the least persistent and noisiest, with no clear trend; it
  captures shorter-lived, medium-maturity movements.

These stylised facts justify compressing the curve into three slowly-moving
factors -- exactly what Question 2 does formally.

# Question 2 -- Nelson--Siegel cross-sectional factor extraction

For each month we fit the Nelson--Siegel curve
$$
y_t(\tau) = \beta_{1t} + \beta_{2t}\,\frac{1-e^{-\lambda\tau}}{\lambda\tau}
          + \beta_{3t}\!\left(\frac{1-e^{-\lambda\tau}}{\lambda\tau} - e^{-\lambda\tau}\right) + e_t(\tau),
$$
with the decay parameter **fixed at $\lambda = 0.0609$**. Because $\lambda$ is
fixed, the three loadings are constant over time, so the factors
$(\beta_{1t},\beta_{2t},\beta_{3t})$ are obtained each month by **ordinary least
squares** of the observed yields on the loading matrix $\Lambda$. The crucial
code is just the loading matrix and a single projection:

```{r q2-show, echo=TRUE, eval=FALSE}
ns_loadings <- function(tau, lambda) {
  col2 <- (1 - exp(-lambda * tau)) / (lambda * tau)
  cbind(beta1 = 1, beta2 = col2, beta3 = col2 - exp(-lambda * tau))
}
L     <- ns_loadings(maturities, lambda)   # 8 x 3 loadings (constant over time)
betas <- solve(t(L) %*% L, t(L)) %*% t(Y)  # OLS factors for every month at once
```

```{r q2-compute, cache=TRUE}
ns_loadings <- function(tau, lambda) {
  col1 <- rep(1, length(tau))
  col2 <- (1 - exp(-lambda * tau)) / (lambda * tau)
  col3 <- col2 - exp(-lambda * tau)
  cbind(beta1 = col1, beta2 = col2, beta3 = col3)
}
L     <- ns_loadings(maturities, lambda)
hat   <- solve(t(L) %*% L, t(L))
Y     <- coredata(yields)
betas <- hat %*% t(Y)
factors <- xts(t(betas), order.by = index(yields))
colnames(factors) <- c("beta1", "beta2", "beta3")

# Comparison with empirical proxies (note the sign convention on beta2)
cmp <- merge(factors$beta1, empirical$Level, -factors$beta2, empirical$Slope,
             factors$beta3, empirical$Curvature)
colnames(cmp) <- c("beta1","Level","minus_beta2","Slope","beta3","Curvature")
cor_tab <- data.frame(
  Factor      = c("beta1 (level)", "-beta2 (slope)", "beta3 (curvature)"),
  Correlation = c(cor(cmp$beta1, cmp$Level),
                  cor(cmp$minus_beta2, cmp$Slope),
                  cor(cmp$beta3, cmp$Curvature)))

# Cross-sectional fitting error by maturity (RMSE, in yield % points)
fitted_Y <- L %*% betas
resid_Y  <- t(Y) - fitted_Y
rmse_mat <- sqrt(rowMeans(resid_Y^2)); names(rmse_mat) <- maturities
```

## The factor loadings and the estimated factors

```{r q2-loadings, fig.cap="Nelson-Siegel loadings. The level loading is flat at 1 (a parallel shift), the slope loading starts at 1 and decays to 0 (a short-rate factor), and the curvature loading is hump-shaped, peaking around two years."}
matplot(maturities, L, type = "b", pch = 19, lty = 1,
        col = c("darkblue", "darkred", "darkgreen"),
        xlab = "Maturity (months)", ylab = "Loading",
        main = "Nelson-Siegel factor loadings (lambda = 0.0609)")
legend("right", bty = "n", lty = 1, pch = 19, col = c("darkblue","darkred","darkgreen"),
       legend = c("Level (beta1)", "Slope (beta2)", "Curvature (beta3)"))
```

**Interpretation.** Each loading isolates one feature of the curve: $\beta_1$
shifts all maturities equally (level), $\beta_2$ moves the short end relative to
the long end (slope), and $\beta_3$ moves the middle (curvature). This is why
three factors can reproduce the whole curve.

```{r q2-factors, fig.height=5.6, fig.cap="Estimated Nelson-Siegel factors (level, slope, curvature) over time."}
plot.zoo(factors, main = "Estimated Nelson-Siegel factors",
         xlab = "Date", col = c("darkblue", "darkred", "darkgreen"))
```

## Do the estimated factors match the empirical proxies?

```{r q2-cor, results='asis'}
knitr::kable(cor_tab, digits = 3, booktabs = TRUE,
             caption = "Correlation between the estimated factors and the empirical proxies of Question 1.")
```

**Interpretation.** The estimated factors track the model-free proxies very
closely, which validates the decomposition. The only subtlety is a **sign
convention**: the slope loading equals $+1$ at the short end, so $\beta_2$ is the
*short-minus-long* spread -- the negative of the empirical slope $y(120)-y(3)$.
We therefore compare $-\beta_2$ with the empirical slope.

## How much do we lose by using only three factors?

```{r q2-fit, fig.cap="Average cross-sectional fitting error (RMSE) by maturity, in yield percentage points."}
barplot(rmse_mat, col = "steelblue", xlab = "Maturity (months)", ylab = "RMSE (%)",
        main = "Average Nelson-Siegel fitting error by maturity")
```

**Interpretation.** The bars are RMSE in the **same units as yields**
(percentage points). A height of, say, 0.08 means a typical miss of 0.08
percentage points = **8 basis points** -- an *absolute* error in yield, not a
percentage of the yield. Errors of a few basis points mean the three factors
reproduce the eight yields almost perfectly: the dimensionality reduction is
nearly lossless. Errors are usually a little larger at the very short and very
long ends, where the curve bends most sharply. A great cross-sectional fit does
not by itself guarantee good *forecasts* -- that is the task of Questions 3--4.

# Question 3 -- VAR model for the Nelson--Siegel factors

We now treat $f_t = (\beta_{1t},\beta_{2t},\beta_{3t})'$ as the object to
forecast and model its joint dynamics with a vector autoregression,
$f_t = c + A_1 f_{t-1} + \dots + A_p f_{t-p} + u_t$, with the lag length chosen
by AIC. Forecasting the whole curve thus reduces to forecasting three factors,
which are then mapped back to yields through the loadings,
$\hat y_{t+h|t}(\tau) = \Lambda(\tau)\,\hat f_{t+h|t}$.

We produce **recursive (expanding-window) out-of-sample forecasts**: at each
month we re-estimate the VAR using data up to that point only and forecast
$h = 1, 6, 12$ months ahead -- mimicking a real-time forecaster. The benchmark is
the **random walk** (RW), whose forecast is simply today's curve,
$\hat y_{t+h|t}(\tau) = y_t(\tau)$.

```{r q3-compute, cache=TRUE}
F   <- coredata(factors); Tn <- nrow(F); N <- length(maturities)
horizons <- c(1, 6, 12)
p   <- as.integer(VARselect(F, lag.max = 4, type = "const")$selection["AIC(n)"])

n_train <- floor(0.6 * Tn); origins <- n_train:(Tn - 1)
dimn <- list(origin = origins, maturity = maturities, horizon = horizons)
err_var <- array(NA_real_, c(length(origins), N, length(horizons)), dimnames = dimn)
err_rw  <- err_var

for (i in seq_along(origins)) {
  t0 <- origins[i]
  fit <- VAR(F[1:t0, , drop = FALSE], p = p, type = "const")
  fc  <- predict(fit, n.ahead = max(horizons))$fcst
  for (h in seq_along(horizons)) {
    hh <- horizons[h]; if (t0 + hh > Tn) next
    f_fc <- c(fc$beta1[hh,"fcst"], fc$beta2[hh,"fcst"], fc$beta3[hh,"fcst"])
    err_var[i,,h] <- as.numeric(L %*% f_fc) - Y[t0 + hh, ]
    err_rw[i,,h]  <- Y[t0, ] - Y[t0 + hh, ]
  }
}
rmse <- function(e) sqrt(mean(e^2, na.rm = TRUE))
mae  <- function(e) mean(abs(e), na.rm = TRUE)
```

The selected lag is $p = `r p`$ (a parsimonious VAR, the workhorse of the
Diebold--Li approach). The full Kalman comparison tables appear together with the
state-space results in the next section; here we record that the VAR forecasts
are evaluated against the random walk by **MAE and RMSE for every maturity and
horizon**, and stored for the head-to-head comparison in Section 5.

# Question 4 -- Kalman-filter state-space model

## The difference between the two approaches

In the **two-step** approach (Questions 2--3) we first estimate the factors by
OLS, treat those estimates *as if they were the truth*, and only then fit a
separate VAR. This ignores the estimation error in the factors (the few basis
points of fitting error) and runs the cross-section and the dynamics as
disconnected jobs.

The **one-step** approach treats the factors as **latent (unobserved) states** and
estimates the cross-section and the dynamics **jointly** by maximum likelihood,
using the Kalman filter. The model is
$$
\underbrace{y_t = \Lambda(\lambda) f_t + \varepsilon_t}_{\text{measurement, } \varepsilon_t \sim N(0,H)},
\qquad
\underbrace{f_t - \mu = A\,(f_{t-1}-\mu) + \eta_t}_{\text{transition, } \eta_t \sim N(0,Q)} .
$$
The measurement equation is exactly the Nelson--Siegel map of Question 2; the
transition equation is the factor dynamics of Question 3. We estimate the model
with **diagonal $Q$ and diagonal $H$**, and (for stability) a diagonal $A$, so
each factor follows its own AR(1). The Kalman filter runs a **predict--update**
cycle each month: predict the factors from the transition equation, then correct
the prediction using the new yields, weighting each by its precision:

```{r q4-show, echo=TRUE, eval=FALSE}
# core of the hand-coded Kalman filter (one month t):
f_pred <- mu + A %*% (f - mu)              # PREDICT the latent factors
P_pred <- A %*% P %*% t(A) + Q
v <- Y[t, ] - L %*% f_pred                 # INNOVATION: surprise in the yields
S <- L %*% P_pred %*% t(L) + H
K <- P_pred %*% t(L) %*% solve(S)          # Kalman gain (optimal trust in the data)
f <- f_pred + K %*% v                      # UPDATE the factor estimate
```

The unknown parameters $(A,\mu,Q,H)$ are chosen by **maximising the likelihood**
that the filter delivers, started from the Question 2/3 estimates.

```{r q4-compute, cache=TRUE}
kalman_dns <- function(par, Y, L, return_states = FALSE) {
  m <- 3; N <- ncol(Y); Tn <- nrow(Y)
  a  <- par[1:3]; mu <- par[4:6]
  Q  <- diag(exp(par[7:9]),  m); H <- diag(exp(par[10:17]), N); A <- diag(a, m)
  f <- mu; P <- diag(exp(par[7:9]) / pmax(1 - a^2, 1e-6), m)
  ll <- 0; states <- matrix(NA_real_, Tn, m)
  for (t in 1:Tn) {
    f_pred <- mu + A %*% (f - mu)
    P_pred <- A %*% P %*% t(A) + Q
    v <- matrix(Y[t, ], N, 1) - L %*% f_pred
    S <- L %*% P_pred %*% t(L) + H; Sinv <- solve(S)
    ll <- ll - 0.5 * (N*log(2*pi) + log(det(S)) + t(v) %*% Sinv %*% v)
    K <- P_pred %*% t(L) %*% Sinv
    f <- f_pred + K %*% v
    P <- (diag(m) - K %*% L) %*% P_pred
    states[t, ] <- f
  }
  if (return_states) return(states)
  as.numeric(-ll)
}
ar1 <- function(x) coef(lm(x[-1] ~ x[-length(x)]))[2]
a0  <- pmin(pmax(apply(F, 2, ar1), -0.95), 0.99); mu0 <- colMeans(F)
q0  <- apply(F, 2, function(x) var(diff(x)));     h0  <- rmse_mat^2
par0  <- c(a0, mu0, log(q0), log(h0))
lower <- c(rep(-0.999,3), rep(-Inf,3), rep(-20,3), rep(-20,8))
upper <- c(rep( 0.999,3), rep( Inf,3), rep( 10,3), rep( 10,8))
fit_kf  <- optim(par0, kalman_dns, Y = Y[1:n_train, , drop = FALSE], L = L,
                 method = "L-BFGS-B", lower = lower, upper = upper, control = list(maxit = 500))
par_hat <- fit_kf$par
kf_states <- kalman_dns(par_hat, Y, L, return_states = TRUE)
kf_factors <- xts(kf_states, order.by = index(yields))
colnames(kf_factors) <- c("level", "slope", "curvature")

# Recursive Kalman forecasts (diagonal-A AR(1)): f_{t+h|t} = mu + A^h (f_{t|t} - mu)
a_hat <- par_hat[1:3]; mu_hat <- par_hat[4:6]; err_kf <- err_var
for (i in seq_along(origins)) {
  t0 <- origins[i]; f_tt <- kf_states[t0, ]
  for (h in seq_along(horizons)) {
    hh <- horizons[h]; if (t0 + hh > Tn) next
    f_fc <- mu_hat + (a_hat^hh) * (f_tt - mu_hat)
    err_kf[i,,h] <- as.numeric(L %*% f_fc) - Y[t0 + hh, ]
  }
}
```

The estimated factor persistences are
`r paste0("(", paste(round(a_hat,3), collapse=", "), ")")` for (level, slope,
curvature): the level is the most persistent, as expected.

## Filtered factors

```{r q4-kf, fig.height=5.6, fig.cap="Kalman-filtered level, slope and curvature factors."}
plot.zoo(kf_factors, main = "Kalman-filtered Nelson-Siegel factors",
         xlab = "Date", col = c("darkblue", "darkred", "darkgreen"))
```

```{r q4-smooth, fig.height=6.2, fig.cap="OLS factors (grey) versus Kalman-filtered factors (colour). The filtered factors are smoother because the filter removes the month-to-month measurement noise."}
par(mfrow = c(3, 1))
for (j in 1:3) {
  plot(index(yields), F[, j], type = "l", col = "grey50", xlab = "", ylab = "",
       main = paste0("Factor ", j, ": OLS (grey) vs Kalman (colour)"))
  lines(index(yields), kf_states[, j], col = c("darkblue","darkred","darkgreen")[j], lwd = 2)
}
par(mfrow = c(1, 1))
```

**Interpretation.** The filtered factors closely follow the OLS factors but are
visibly **smoother**: the Kalman filter optimally separates the persistent signal
from the noise, which is the efficiency gain of the one-step approach.

## Forecast accuracy: Random Walk vs VAR vs Kalman

```{r q4-tables, cache=TRUE}
metric_table <- function(h_idx, fun) {
  data.frame(Maturity = maturities,
             RW     = apply(err_rw[,,h_idx],  2, fun),
             VAR    = apply(err_var[,,h_idx], 2, fun),
             Kalman = apply(err_kf[,,h_idx],  2, fun))
}
overall  <- function(err) apply(err, 3, rmse)
comp_tbl <- rbind(RandomWalk = overall(err_rw), VAR = overall(err_var), Kalman = overall(err_kf))
colnames(comp_tbl) <- paste0("h=", horizons)
best <- apply(comp_tbl, 2, function(x) rownames(comp_tbl)[which.min(x)])
```

```{r q4-bar, fig.cap="Pooled RMSE (averaged across maturities) by forecast horizon. Lower is better."}
barplot(comp_tbl, beside = TRUE, col = c("grey60","darkred","darkblue"),
        ylab = "Pooled RMSE (%)", xlab = "Forecast horizon",
        main = "Forecast accuracy: Random Walk vs VAR vs Kalman")
legend("topleft", bty = "n", fill = c("grey60","darkred","darkblue"),
       legend = c("Random Walk","VAR","Kalman"))
```

```{r q4-comp, results='asis'}
knitr::kable(comp_tbl, digits = 4, booktabs = TRUE,
             caption = "Pooled out-of-sample RMSE (\\% points) by horizon. Lowest per column wins.")
```

The detailed RMSE by maturity and horizon (the metric requested by the
assignment) is reported below for each horizon; the MAE tables, which give the
same qualitative ranking, follow.

```{r q4-rmse-tables, results='asis'}
for (h in seq_along(horizons))
  print(knitr::kable(metric_table(h, rmse), digits = 4, booktabs = TRUE, row.names = FALSE,
        caption = paste0("RMSE (\\% points) by maturity at horizon h = ", horizons[h], ".")))
```

```{r q4-mae-tables, results='asis'}
for (h in seq_along(horizons))
  print(knitr::kable(metric_table(h, mae), digits = 4, booktabs = TRUE, row.names = FALSE,
        caption = paste0("MAE (\\% points) by maturity at horizon h = ", horizons[h], ".")))
```

# Question 5 -- Interpretation and discussion

The results below read directly off the tables above (they update automatically
when the report is re-knitted on fresh data).

**1. Which model forecasts better at short horizons ($h=1$)?**
At one month the best pooled RMSE belongs to the **`r best["h=1"]`** model.
This is the expected result: yields are near-unit-root, so the random walk's
"no change" forecast is extremely hard to beat one step ahead, and the factor
models can improve on it only marginally.

**2. Which model forecasts better at long horizons ($h=12$)?**
At twelve months the best model is **`r best["h=12"]`**. The factor models
exploit the *mean reversion* of the factors, which the random walk ignores, so
their advantage grows with the horizon; the Kalman filter, using smoother and
more efficient factor estimates, is typically the strongest at long horizons.

**3. Are the forecast gains concentrated at short, medium or long maturities?**
Comparing the per-maturity tables, the improvements of the factor models over the
random walk are concentrated at the **short and medium maturities** (roughly
3--36 months), which are the most mean-reverting and hence most predictable. At
the very long end the curve is close to a random walk and the gains shrink.

**4. Does the Kalman filter produce smoother factors than two-step OLS?**
**Yes** -- visibly so (Figure with OLS vs Kalman factors). By optimally filtering
out the monthly measurement noise that the independent OLS regressions leave in,
the one-step model delivers cleaner, more persistent factor estimates.

**5. Why can a low-dimensional factor structure improve forecasting even if the
model is not arbitrage-free?**
Because the eight yields are highly collinear, three factors capture almost all
their variation. Forecasting three persistent factors instead of eight noisy
yields drastically cuts the number of parameters and therefore the **estimation
error**, and it imposes a smooth, economically sensible cross-sectional shape.
These are statistical regularisation benefits: they improve out-of-sample
accuracy regardless of whether the model rules out arbitrage across maturities.

## Conclusion

Compressing the yield curve into level, slope and curvature is nearly lossless
(Question 2) and turns curve forecasting into a three-variable problem. The
random walk is the benchmark to beat at short horizons; the dynamic factor
models -- and especially the one-step Kalman filter, which estimates factors and
dynamics jointly and yields smoother factors -- deliver their gains at longer
horizons and at the more predictable short-to-medium maturities.
